
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{float}
\usepackage{cleveref}
\usepackage{hyperref}
\usepackage{csquotes}
\usepackage{multicol}
\usepackage{booktabs}

\begin{document}

\title{MACM 101 Chapter 2 Homework - Halting Problem}
\author{Alexander Ng}
\date{October 11, 2024}

\maketitle

Assume $H$ is a Turing machine that solves the halting problem.

This function takes an input $P$, which is a program, and $x$, which is an
input to the program, and returns $\mathbf{T}$ if $P$ halts on input $x$, and 
$\mathbf{F}$ otherwise.

\begin{equation*}
  H(P, x) = \begin{cases}
    P \text{ halts on } x & \mathbf{T} \\
    P \text{ does not halt on } x & \mathbf{F}
  \end{cases}
\end{equation*}

Let $D$ be another Turing machine that takes an input $P$, which is a program,
and returns $H(P, P)$. That is, $D$ takes a program $P$ as input and returns
$\mathbf{T}$ if $P$ halts on $P$, and $\mathbf{F}$ otherwise.

\begin{equation*}
  D(P) = \begin{cases}
    H(P, P) & \text{ D never halts} \\
    & \text{ D halts immediately}
  \end{cases}
\end{equation*}

Consider $D(D)$:

\begin{equation*}
  D(D) = \begin{cases}
    \text{If } H(D, D) = \mathbf{T}, \text{ then } D(D) \text{ goes into an infinite loop.} \\
    \text{If } H(D, D) = \mathbf{F}, \text{ then } D(D) \text{ halts immediately.}
  \end{cases}
\end{equation*}

Now:

\begin{itemize}
  \item If $H(D, D) = \mathbf{T}$, this implies $D(D)$ should halt, but by the definition of $D$, if $H(D, D) = \mathbf{T}$, $D(D)$ enters an infinite loop, which is a contradiction.
  \item If $H(D, D) = \mathbf{F}$, this implies $D(D)$ does not halt, but by the definition of $D$, if $H(D, D) = \mathbf{F}$, $D(D)$ halts immediately, which is also a contradiction.
\end{itemize}

Since both cases lead to contradictions, the assumption that $H$ exists is false.

Therefore, the halting problem is unsolvable.

$\qed$

\end{document}
